---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
##Summary of the assignment

This assignment makes use of data from a personal activity monitoring device, such as Fitbit, Nike Fuelband, or Jawbone Up, which collects data at 5 minute intervals throughout the day. The data consists of two months of data from an anonymous individual collected during the months of October and November 2012, and include the number of steps taken in 5 minute intervals each day.

The variables included in this dataset are:

- steps: Number of steps taking in a 5-minute interval (missing values are coded as NA)
- date: The date on which the measurement was taken in YYYY-MM-DD format
- interval: Identifier for the 5-minute interval in which measurement was taken

The dataset is stored in a CSV file and there are a total of 17,568 observations in this dataset.


The **goal** of this assignment is to write a report that answers the following questions:

1. Loading and preprocessing the data
2. What is mean total number of steps taken per day?
3. What is the average daily activity pattern?
4. Imputing missing values
5. Are there differences in activity patterns between weekdays and weekends?

## Loading and preprocessing the data

First, the CSV file that contains 17,568 observations will be loaded by using *read.csv()*.

```{r}
activity_data <- read.csv("activity.csv", head = TRUE, sep = ",")
```

Before I continue with the assignment, I want to find out more about the contents of the CSV file. 

First, I will check the names of columns by using *names()*.

```{r, echo=TRUE}
names(activity_data)
```

Then I will display the internal structure of the activity_data object by using *str()*.

```{r, echo=TRUE}
str(activity_data)
```

What I can see from the structure displayed above, is that data belonging to the *steps* column are class integer (note that I could also check this by calling *class(activity_data$steps)*). I also know that there will be some missing values (NA) in this dataset. 

```{r, echo=TRUE}
class(activity_data$steps)
```

I can also see that dates are class factor and written in the following format: YYYY-MM-DD.

Interval holds integers and no NA values.

Now that I know how my data looks like, I can continue with the assignment.

## What is mean total number of steps taken per day?

For this part of the assignment, I will first calculate the total number of steps taken per day by using *aggregate()* and ignore the missing values (NA) in the dataset by using *na.omit()*.

```{r, echo=TRUE}
 steps_per_day <- na.omit(aggregate(activity_data$steps, list(date = activity_data$date), FUN = sum))
```

Once I run the code, I can see how many steps a person has made each day. Below you can see a total number of steps (*x*) for the first five days: 


```{r, echo=TRUE}
 head(steps_per_day, 5)
```

Since for the 2012-10-01 we had a not available (NA) number of steps for all intervals, the whole day is skipped from the calculation. I cannot assume that a person made 0 steps, so I will not assign 0 to day 1 (2012-10-01). The following day (2012-10-02), the person made 117 steps in the interval 2210 and 9 steps in the interval 2215, which sums up to (in total) **126** steps taken on the day 2012-10-02.


Then I will make a histogram of the total number of steps taken each day.

```{r, echo=TRUE}
barplot(steps_per_day$x, 
        main="Steps per day", 
        names.arg = steps_per_day$date,
        xlab="Date")
```

To calculate the mean and median total number of steps per day, I will need the following 2 R functions:

- *mean()*
- *median()*

I will find the mean of the number of steps taken daily:

```{r, echo=TRUE}
mean(steps_per_day$x)
```

I will do the same to find the median:
```{r, echo=TRUE}
median(steps_per_day$x)
```


## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
